Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish

### 2024-10-01 manuscript.tex
reworked figures 3, 7, 8, 9 and 10. minor adjustments to text. added scripts to load interfaces and recreate plots

The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects. Below this, users can save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}F). New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}G), and users can select different metrics for some plot types e.g. heatmaps (Figure \ref{fig:GUI_overview}H).

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{GUI_overview.png}
    \caption{Overview of the GUI. A: Sidebar controls; B: Display area; C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Save and load buttons; G: Interface for starting new tours; H: Metric selection interface.}
    \label{fig:GUI_overview}
\end{figure}

@@

where $c_{ij}, i=1, ..., p$ (number of features); $j=1, ..., k$ (number of clusters) are a summary of each feature in each cluster. Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$. %For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection. The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data. For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$. The first is the overall fraction, where counts are normalised by the overall number of observations $n$. It is generally not particularly useful. The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster. The last, $f_f$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters \textcolor{red}{Not sure this last explanation is good - at least I don't understand what I've written}

@@

\textcolor{red}{It might be better to standardise these intra fractions. For A, one should divide by cluster count, and then compute activity fraction. For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster. For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors. The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F). We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6}
    \label{fig:winter_activ_cluster_highlights}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Tourists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
    \label{fig:winter_cl7_init}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing'' and ``using health facilities'' reveals the preferences and overlap of clusters 1 and 3.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
    \label{fig:winter_cl7_post}
\end{figure}

>>>>

The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects. Below this, there are three additional interfaces (Figure \ref{fig:GUI_overview}F). The first can be used to adjust the number of bins in histograms—more bins result in higher resolution but slower display updates. The second allows users to animate tours, automatically shifting to the next projection after a specified amount of time. The third offers the ability to hide projection axes with a norm smaller than a chosen threshold, helping to reduce clutter. Users can also save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}G). New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g. heatmaps (Figure \ref{fig:GUI_overview}I).

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{GUI_overview.png}
    \caption{Overview of the GUI. A: Sidebar controls; B: Display area; C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Interfaces for adjusting the number of bins of histograms, animating tours and blending out projection axes with a low norm; G: Save and load buttons; H: Interface for starting new tours; I: Metric selection interface.}
    \label{fig:GUI_overview}
\end{figure}

@@

where $c_{ij}, i=1, ..., p$ (number of features); $j=1, ..., k$ (number of clusters) are a summary of each feature in each cluster. Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$. %For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection. The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data. For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$. The first is the overall fraction, where counts are normalised by the overall number of observations $n$. It gives a quick overview on the overall magnitude of the features. The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster. The last, $f^{f}$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters.

@@

In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster. For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F). Some datapoints appear to be highlighted always, which occurs due to overlap of many datapoints in one spot. We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6.}
    \label{fig:winter_activ_cluster_highlights}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Tourists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
    \label{fig:winter_cl7_init}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing'' and ``using health facilities'' reveals the preferences and overlap of clusters 1 and 3.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected. Projection axes with a norm < 0.1 have been blend out to reduce cluttering. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
    \label{fig:winter_cl7_post}
\end{figure}
